# Alpha Diversity

The alpha diversity is the mean diversity of a species in a site at local scale.

It can be calculated using different divesity indices like richness, Shannon index and Simpson index.

## Alpha diversity plot

Plot Alpha diversity using the three different metrics.

```{r, eval = F}
# By Samples
plot_richness(ps_rare, x = "samples", measures = c("Observed", "Shannon", "Simpson"))
```

It is possible to separate by different categories than only Samples

```{r, eval = F}
# By Islands
plot_richness(ps_rare, x = "Islands", measures = c("Observed", "Shannon", "Simpson", color = "Island"))
```

```{r, eval = F }
# By Communities
plot_richness(ps_rare, x = "Communities", measures = c("Obserced", "Shannon", "Simpson"), color = "Communities")
```


## Box-plots ASV richness

We can also represent it in Box-plots

```{r, eval = F}
Boxplot <- plot_richness(ps_rare, x = "sample", measures = "Observed")

ggplot(Boxplot$data, aes(x = sample , y = value, fill = sample)) + 
  geom_violin() +
  geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) + 
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Richness") + xlab("Island") + theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank()) +
  #scale_fill_manual(values = color_list[c(6,4,8)]) + 
  #scale_color_manual(values = color_list)
```


