# Beta Diversity

The beta diversity is the ratio between regional and local species.

There is multiple way to represent it but a PCA plot is one of the most used way.

Obviously multiple Metrics can be used like Bray-Curtis, Aitchinson, Jaccard, etc..

```{r}
# Library
library("plyr"); packageVersion("plyr")
```


## Metric conversion and different Method computation

```{r, eval = F}
# Compute all the methods in bray-curtis
dist = "bray"
ord_meths = c("NMDS", "PCoA", "MDS", "CCA", "DCA", "RDA")
plist = llply(as.list(ord_meths), function(i, physeq, dist){
        ordi = ordinate(physeq, method=i, distance=dist)
        plot_ordination(physeq, ordi, "samples", color="Island")
}, ps_Norm, dist)

names(plist) <- ord_meths

# Create a dataframe containing each method

pdataframe = ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("PC1", "PC2")
    return(cbind(df, x$data))
})
names(pdataframe)[1] = "method"
```

## PCA plot

We can then plot the different methods togheter or separately.

```{r, eval = F}
# Togheter 
p = ggplot(pdataframe, aes(PC1, PC2, color= Island))
p = p + geom_point(size=2) #geom™_polygon()
p = p + facet_wrap(~method, scales="free")
p = p + scale_fill_brewer(type="qual", palette="Set1")
p = p + scale_colour_brewer(type="qual", palette="Set1")
p
```

## PCA Plot Using Bray-Curtis (Islands)

```{r}
col_gir <- c("#a0be25","#25be4b")

names(col_gir) <- c("Bahamas", "Kauai")

ord_ps = ordinate(ps_Norm, "PCoA", "bray")
g <- plot_ordination(ps_Norm, ord_ps, justDF=T)
head(g)

# extract % variation
tempx <- as.character(round(ord_ps$values$Relative_eig[1],3) *100)
tempy <- as.character(round(ord_ps$values$Relative_eig[2],3) *100)
set.seed(386)

# pcoa plot Giraffe Species
ggplot(g, aes(x=Axis.1, y=Axis.2, color=Island, fill=Island)) + theme_bw() + 
  stat_ellipse(geom = "polygon", type="t", alpha=0.2, level=0.9)  +  # confidence interval
  geom_point(color="black", shape=21, size = 2.5, stroke=0.3) +
  xlab(paste0("PC1 (",tempx,"%)")) + ylab(paste0("PC2 (",tempy,"%)")) +
  theme(axis.title.x = element_text(size=13, margin = margin(t=3)),
        axis.title.y = element_text(size=13, margin = margin(r=2)),
        axis.text.y  = element_text(size=11, color = "black"), 
        axis.text.x  = element_text(size=11, color = "black"),
        strip.text   = element_text(size=13)) +
  theme(panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "grey29")) +
  theme(legend.title = element_text(size = 13), 
        legend.text  = element_text(size = 13),
        legend.key.size = unit(0.6, "cm"),
        legend.position = "right") +
  scale_color_manual(values = col_gir) +
  scale_fill_manual(values  = col_gir)
```

## PCA Plot Using Bray-Curtis (Sex)

```{r}
col_gir <- c("#a0be25","#25be4b")

names(col_gir) <- c("Male", "Female")

ord_ps = ordinate(ps_Norm, "PCoA", "bray")
g <- plot_ordination(ps_Norm, ord_ps, justDF=T)
head(g)

# extract % variation
tempx <- as.character(round(ord_ps$values$Relative_eig[1],3) *100)
tempy <- as.character(round(ord_ps$values$Relative_eig[2],3) *100)
set.seed(386)

# pcoa plot Giraffe Species
ggplot(g, aes(x=Axis.1, y=Axis.2, color=Sex, fill=Sex)) + theme_bw() + 
  stat_ellipse(geom = "polygon", type="t", alpha=0.2, level=0.9)  +  # confidence interval
  geom_point(color="black", shape=21, size = 2.5, stroke=0.3) +
  xlab(paste0("PC1 (",tempx,"%)")) + ylab(paste0("PC2 (",tempy,"%)")) +
  theme(axis.title.x = element_text(size=13, margin = margin(t=3)),
        axis.title.y = element_text(size=13, margin = margin(r=2)),
        axis.text.y  = element_text(size=11, color = "black"), 
        axis.text.x  = element_text(size=11, color = "black"),
        strip.text   = element_text(size=13)) +
  theme(panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "grey29")) +
  theme(legend.title = element_text(size = 13), 
        legend.text  = element_text(size = 13),
        legend.key.size = unit(0.6, "cm"),
        legend.position = "right") +
  scale_color_manual(values = col_gir) +
  scale_fill_manual(values  = col_gir)
```

## PCA Plot Using Bray-Curtis (Sample Type)

```{r}
col_gir <- c("#a0be25","#25be4b")

names(col_gir) <- c("Cecum", "Scrape")

ord_ps = ordinate(ps_Norm, "PCoA", "bray")
g <- plot_ordination(ps_Norm, ord_ps, justDF=T)
head(g)

# extract % variation
tempx <- as.character(round(ord_ps$values$Relative_eig[1],3) *100)
tempy <- as.character(round(ord_ps$values$Relative_eig[2],3) *100)
set.seed(386)

# pcoa plot
ggplot(g, aes(x=Axis.1, y=Axis.2, color=Sample.Type, fill=Sample.Type)) + theme_bw() + 
  stat_ellipse(geom = "polygon", type="t", alpha=0.2, level=0.9)  +  # confidence interval
  geom_point(color="black", shape=21, size = 2.5, stroke=0.3) +
  xlab(paste0("PC1 (",tempx,"%)")) + ylab(paste0("PC2 (",tempy,"%)")) +
  theme(axis.title.x = element_text(size=13, margin = margin(t=3)),
        axis.title.y = element_text(size=13, margin = margin(r=2)),
        axis.text.y  = element_text(size=11, color = "black"), 
        axis.text.x  = element_text(size=11, color = "black"),
        strip.text   = element_text(size=13)) +
  theme(panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "grey29")) +
  theme(legend.title = element_text(size = 13), 
        legend.text  = element_text(size = 13),
        legend.key.size = unit(0.6, "cm"),
        legend.position = "right") +
  scale_color_manual(values = col_gir) +
  scale_fill_manual(values  = col_gir)
```

## PCA Plot using wheighted Unifrac (Island)

```{r}
col_gir <- c("#a0be25","#25be4b")

names(col_gir) <- c("Bahamas", "Kauai")

ord_ps = ordinate(ps_Norm, "PCoA", "unifrac", weighted = T)
g <- plot_ordination(ps_Norm, ord_ps, justDF=T)
head(g)

# extract % variation
tempx <- as.character(round(ord_ps$values$Relative_eig[1],3) *100)
tempy <- as.character(round(ord_ps$values$Relative_eig[2],3) *100)
set.seed(386)

# pcoa plot Giraffe Species
ggplot(g, aes(x=Axis.1, y=Axis.2, color=Island, fill=Island)) + theme_bw() + 
  stat_ellipse(geom = "polygon", type="t", alpha=0.2, level=0.9)  +  # confidence interval
  geom_point(color="black", shape=21, size = 2.5, stroke=0.3) +
  xlab(paste0("PC1 (",tempx,"%)")) + ylab(paste0("PC2 (",tempy,"%)")) +
  theme(axis.title.x = element_text(size=13, margin = margin(t=3)),
        axis.title.y = element_text(size=13, margin = margin(r=2)),
        axis.text.y  = element_text(size=11, color = "black"), 
        axis.text.x  = element_text(size=11, color = "black"),
        strip.text   = element_text(size=13)) +
  theme(panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = "grey29")) +
  theme(legend.title = element_text(size = 13), 
        legend.text  = element_text(size = 13),
        legend.key.size = unit(0.6, "cm"),
        legend.position = "right") +
  scale_color_manual(values = col_gir) +
  scale_fill_manual(values  = col_gir)
```
