# Beta Diversity

The beta diversity is the ratio between regional and local species.

There is multiple way to represent it but a PCA plot is one of the most used way.

Obviously multiple Metrics can be used like Bray-Curtis, Aitchinson, Jaccard, etc..

```{r}
# Library
library("plyr"); packageVersion("plyr")
```


## Metric conversion and different Method computation

```{r, eval = F}
# Compute all the methods in bray-curtis
dist = "jaccard"
ord_meths = c("NMDS", "PCoA", "MDS")
plist = llply(as.list(ord_meths), function(i, physeq, dist){
        ordi = ordinate(physeq, method=i, distance=dist)
        plot_ordination(physeq, ordi, "samples")
}, ps_Norm, dist)

names(plist) <- ord_meths

# Create a dataframe containing each method

pdataframe = ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("PC1", "PC2")
    return(cbind(df, x$data))
})
names(pdataframe)[1] = "method"
```

## PCA plot

We can then plot the different methods togheter or separately.

```{r, eval = F}
# Togheter 
p = ggplot(pdataframe, aes(PC1, PC2))
p = p + geom_point(size=2) #geom_polygon()
p = p + facet_wrap(~method, scales="free")
p = p + scale_fill_brewer(type="qual", palette="Set1")
p = p + scale_colour_brewer(type="qual", palette="Set1")
p
```

Finish more plots