# Statistical test

## Load Workspace

```{r, eval = F}
load("/Users/matiasbeckerburgos/Master in Bioinformatics /Master Thesis /Hawaiian_feral_Chicken_Microbiome/Backup_workspace/Decomtamination_pipeline.RData")
```

There is multiple statistical test which can be realised:

## Alpha diversity test

### Check normal distribution

By doing an histogram, a Q-Q plot or a shapiro test. You can check for normal distribution.

```{r, eval = F}
hist()
qqplot()
qqline()
shapiro.test()
```

Usually microbiome data doesn't really follow normal distribution so non-parametric test are preferred.

### Exctract variables

```{r}
rich <- estimate_richness(ps_rare)

alpha_div <- sample_data(ps_rare)

alpha_div$Richness <- rich$Observed

alpha_div <- as.matrix(alpha_div)
```

### Kurskal-Wallice test

The Kurskal-Wallice test is a non-parametric test for testing if samples originate from the same distribution. 
 
It is used to compare two or more independant samples of equal or different sample size. 

**significant results**: Indicates that at least one sample stochastically dominates one other sample (but doesn't say which one).

```{r, eval = F}
kw <- kruskal.test(Richness ~ Island, data = alpha_div)
kw
```

It is a good test to start with.

## Beta diversity test

#### Generate distance metrics

```{r, eval = F}
# Set the distance metrics for the matrix
ps_dist_matrix <- phyloseq::distance(ps_Norm, method = "bray")
```

### Adonis test

The Adonis test is a version of a PERMANOVA. It is essentially a test for difference in the between beta diversity in the sample variables.

**Significant result**: Indicates difference between samples variables.

```{r, eval = F}
# Do the Adonis test based on one variables you want
vegan::adonis2(ps_dist_matrix ~ phyloseq::sample_data(ps_Norm)$Sample.Type, permutations = 999)
```

```{r, eval = F}
# The Adonis test based on two variables
vegan::adonis2(ps_dist_matrix ~ phyloseq::sample_data(ps_Norm)$Island + phyloseq::sample_data(ps_Norm)$Sex, permutations = 999, by = "terms")
```

```{r, eval = F}
# The Adonis test based on interation of two variables
vegan::adonis2(ps_dist_matrix ~ phyloseq::sample_data(ps_Norm)$Island * phyloseq::sample_data(ps_Norm)$Sex, permutations = 999, by = "terms")
```

# Save the data

```{r, eval = F}
save.image(file.path(save, "Decomtamination_pipeline.RData"))
```