# Alpha Diversity

The alpha diversity is the mean diversity of a species in a site at local scale.

It can be calculated using different divesity indices like richness, Shannon index and Simpson index.

## Library
```{r}
library(phyloseq)
```

## Load the Backup workspace
```{r}
# Object containing all the ps object
load("/Users/matiasbeckerburgos/Desktop/Master_in_Bioinformatics/Master_Thesis/Hawaiian_feral_Chicken_Microbiome/Backup_workspace/Decomtamination_pipeline_RUN1.RData")
```

## Set the color palette
```{r}
col_gir <- c("#a0be25", "#25be4b")
```

## Alpha diversity plot

Plot Alpha diversity using the three different metrics.

```{r, eval = F}
# By Samples
plot_richness(ps_rare, x = "samples", measures = c("Observed", "Shannon", "Simpson"))
```

It is possible to separate by different categories than only Samples

```{r, eval = F}
# By Islands
plot_richness(ps_rare, x = "Island", measures = c("Observed", "Shannon", "Simpson", color = "Island"))
```
```{r}
# By Sex
plot_richness(ps_rare, x = "Sex", measures = c("Observed", "Shannon", "Simpson"), color = "Sex")
```

```{r}
# By Sample Type
plot_richness(ps_rare, x = "Sample.Type", measures = c("Observed", "Shannon", "Simpson"), color = "Sample.Type")
```

## ASV richness

### Metrics Used

```{r}
Boxplot <- plot_richness(ps_rare, measures = "Chao1")
```

### Plot for different variable
```{r, eval = F}
# Boxplot By Island
ggplot(Boxplot$data, aes(x = Island , y = value, fill = Island)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Richness") + xlab("Island") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Sex
ggplot(Boxplot$data, aes(x = calc_sex , y = value, fill = calc_sex)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Richness") + xlab("Sex") + labs(fill = "Sex") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Sample Type
ggplot(Boxplot$data, aes(x = Sample.Type , y = value, fill = Sample.Type)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Richness") + xlab("Sample Type") + labs(fill = "Sample Type") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Age category
ggplot(Boxplot$data, aes(x = Age_category , y = value, fill = Age_category)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Richness") + xlab("Age Category") + labs(fill = "Age Category") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Location Sampling
ggplot(Boxplot$data, aes(x = location.numeric , y = value, fill = location.numeric)) +
  #geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  #geom_boxplot(width= 0.2, fill = "white") +
  geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Richness") + xlab("Collection Location") + labs(fill = "Collection Location") +
  facet_wrap(~Island, scales = "free") +
  #scale_fill_manual(values = col_gir) + 
  #scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) +
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Scatter plot for Body Mass
ggplot(Boxplot$data, aes(x = Body.mass..g. , y = value)) +
  geom_point(alpha = 1, size = 1.25, color = col_gir[2], pch = 21) +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  ylab("Richness") + xlab("Body mass (g)") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

## Alpha-diversity Richness by removing the outliers

```{r}
# Identify the 5 outliers
rich <- estimate_richness(ps_rare)
alpha_div <- sample_data(ps_rare)
alpha_div <- cbind(alpha_div, rich)

# Create a new ps rare object without the outliers (n= 5)
ps_rare_wo_outliers <- subset_samples(ps_rare, ID %in% c(alpha_div$ID[alpha_div$Simpson >= 0.85]))
```

### Metrics Used
```{r}
Boxplot <- plot_richness(ps_rare_wo_outliers, measures = "Chao1")
```

### Plots for different variable

```{r, eval = F}
# Boxplot By Island
ggplot(Boxplot$data, aes(x = Island , y = value, fill = Island)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Richness") + xlab("Island") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Sex
ggplot(Boxplot$data, aes(x = calc_sex , y = value, fill = calc_sex)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Richness") + xlab("Sex") + labs(fill = "Sex") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Sample Type
ggplot(Boxplot$data, aes(x = Sample.Type , y = value, fill = Sample.Type)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Richness") + xlab("Sample Type") + labs(fill = "Sample Type") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Age category
ggplot(Boxplot$data, aes(x = Age_category , y = value, fill = Age_category)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Richness") + xlab("Age Category") + labs(fill = "Age Category") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Location Sampling
ggplot(Boxplot$data, aes(x = location.numeric , y = value, fill = location.numeric)) +
  #geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  #geom_boxplot(width= 0.2, fill = "white") +
  geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Richness") + xlab("Collection Location") + labs(fill = "Collection Location") +
  facet_wrap(~Island, scales = "free") +
  #scale_fill_manual(values = col_gir) + 
  #scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) +
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Scatter plot for Body Mass
ggplot(Boxplot$data, aes(x = Body.mass..g. , y = value)) +
  geom_point(alpha = 1, size = 1.25, color = col_gir[2], pch = 21) +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  ylab("Richness") + xlab("Body mass (g)") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

## Alpha-diversity eveness

### Metrics used

```{r}
library("microbiome")
p <- microbiome::evenness(otu_table(ps_rare))
p

# Create Matrix with eveness value
evenness_Box <- sample_data(ps_rare)

evenness_Box$pielou <- p$pielou

evenness_div <- as.matrix(evenness_Box)
```

### Plots for different variable

```{r, eval = F}
# Boxplot By Island
ggplot(evenness_Box, aes(x = Island , y = pielou, fill = Island)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Evenness") + xlab("Island") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())

```

```{r}
# Boxplot By Sex
ggplot(evenness_Box, aes(x = calc_sex , y = pielou, fill = calc_sex)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Evenness") + xlab("Island") + labs(fill = "Sex") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Sample Type
ggplot(evenness_Box, aes(x = Sample.Type , y = pielou, fill = Sample.Type)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Evenness") + xlab("Island") + labs(fill = "Sample Type") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Age category
ggplot(evenness_Box, aes(x = Age_category , y = pielou, fill = Age_category)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Evenness") + xlab("Age Category") + labs(fill = "Age Category") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Location Sampling
ggplot(evenness_Box, aes(x = location.numeric , y = pielou, fill = location.numeric)) +
  #geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  #geom_boxplot(width= 0.2, fill = "white") +
  geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Evenness") + xlab("Collection Location") + labs(fill = "Collection Location") +
  facet_wrap(~Island, scales = "free") +
  #scale_fill_manual(values = col_gir) + 
  #scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) +
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Scatter plot for Body Mass
ggplot(evenness_Box, aes(x = Body.mass..g. , y = pielou)) +
  geom_point(alpha = 1, size = 1.25, color = col_gir[2], pch = 21) +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  ylab("Evenness") + xlab("Body mass (g)") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

## Alpha-diversity eveness by removing the outliers

### Metrics used

```{r}
library("microbiome")
p <- microbiome::evenness(otu_table(ps_rare_wo_outliers))
p

# Create Matrix with eveness value
evenness_Box <- sample_data(ps_rare_wo_outliers)

evenness_Box$pielou <- p$pielou

evenness_div <- as.matrix(evenness_Box)
```

### Plots for different variable

```{r, eval = F}
# Boxplot By Island
ggplot(evenness_Box, aes(x = Island , y = pielou, fill = Island)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Evenness") + xlab("Island") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())

```

```{r}
# Boxplot By Sex
ggplot(evenness_Box, aes(x = calc_sex , y = pielou, fill = calc_sex)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Evenness") + xlab("Island") + labs(fill = "Sex") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Sample Type
ggplot(evenness_Box, aes(x = Sample.Type , y = pielou, fill = Sample.Type)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Evenness") + xlab("Island") + labs(fill = "Sample Type") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Age category
ggplot(evenness_Box, aes(x = Age_category , y = pielou, fill = Age_category)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Evenness") + xlab("Age Category") + labs(fill = "Age Category") +
  scale_fill_manual(values = col_gir) + 
  scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Boxplot By Location Sampling
ggplot(evenness_Box, aes(x = location.numeric , y = pielou, fill = location.numeric)) +
  #geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  #geom_boxplot(width= 0.2, fill = "white") +
  geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("Evenness") + xlab("Collection Location") + labs(fill = "Collection Location") +
  facet_wrap(~Island, scales = "free") +
  #scale_fill_manual(values = col_gir) + 
  #scale_color_manual(values = col_gir) +
  theme_bw() + 
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) +
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

```{r}
# Scatter plot for Body Mass
ggplot(evenness_Box, aes(x = Body.mass..g. , y = pielou)) +
  geom_point(alpha = 1, size = 1.25, color = col_gir[2], pch = 21) +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  ylab("Evenness") + xlab("Body mass (g)") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank())
```

## Alpha-diversity faith phylogenetic diversity (look if use or not)

```{r}
library("picante")

p2 <- picante::pd(samp = as.data.frame(otu_table(ps_rare)),
                  tree = phy_tree(ps_rare), include.root = FALSE)

# Create Matrix with eveness value
Faith_Box <- sample_data(ps_rare)

Faith_Box$PD <- p2$PD

Faith_div <- as.matrix(Faith_Box)
```

```{r}
# Boxplot By Island

ggplot(Faith_Box, aes(x = Island , y = PD, fill = Island)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("PD") + xlab("Island") 
  #theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank()) +
  #scale_fill_manual(values = color_list[c(6,4,8)]) + 
  #scale_color_manual(values = color_list)
```

```{r}
# Boxplot By Sex


ggplot(Faith_Box, aes(x = Sex , y = PD, fill = Sex)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("PD") + xlab("Sex") 
  #theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank()) +
  #scale_fill_manual(values = color_list[c(6,4,8)]) + 
  #scale_color_manual(values = color_list)
```

```{r}
# Boxplot By Sample Type


ggplot(Faith_Box, aes(x = Sample.Type , y = PD, fill = Sample.Type)) +
  geom_violin(alpha = 1, color = "black", lwd= 0.36, width = 0.7) +
  geom_boxplot(width= 0.2, fill = "white") +
 #geom_boxplot(alpha = 1, color = "black", outlier.color = "white", outlier.size = 0, lwd= 0.36, width = 0.7) +
  geom_point(position = position_jitterdodge(jitter.width = 0.35), alpha = 1, size = 1.25, color = "black", pch = 21) +
  ylab("PD") + xlab("Sample Type") 
  #theme_bw() + 
  theme(axis.title.x = element_text(size=18, margin=margin(t=7)), axis.title.y = element_text(size=18, margin=margin(r=8)), axis.text.y = element_text(size=18, color = "black"), axis.text.x = element_text(size=18, color="black"), strip.text.x = element_blank()) + 
  theme(panel.border = element_rect(colour = "grey29"), panel.grid.minor = element_blank(), panel.grid.major = element_blank()) +
  #scale_fill_manual(values = color_list[c(6,4,8)]) + 
  #scale_color_manual(values = color_list)

```