# Statistical test

## Load Workspace

```{r, eval = F}
load("/Users/matiasbeckerburgos/Desktop/Master_in_Bioinformatics/Master_Thesis/Hawaiian_feral_Chicken_Microbiome/Backup_workspace/Decomtamination_pipeline_RUN1.RData")
```

There is multiple statistical test which can be realised:

## Alpha diversity test

### Check normal distribution

By doing an histogram, a Q-Q plot or a shapiro test. You can check for normal distribution.

```{r, eval = F}
hist()
qqplot()
qqline()
shapiro.test()
```

Usually microbiome data doesn't really follow normal distribution so non-parametric test are preferred.

### Exctract variables

```{r}
# For Richness
rich <- estimate_richness(ps_rare)

alpha_div <- sample_data(ps_rare)

alpha_div <- cbind(alpha_div, rich)

alpha_div_matrix <- as.matrix(alpha_div)
```

### Kurskal-Wallice test

The Kurskal-Wallice test is a non-parametric test for testing if samples originate from the same distribution. 
 
It is used to compare two or more independant samples of equal or different sample size. 

**significant results**: Indicates that at least one sample stochastically dominates one other sample (but doesn't say which one).

```{r, eval = F}
kw <- kruskal.test(Simpson ~ Age_category, data = alpha_div_matrix)
kw
```

It is a good test to start with.

#### Variable with significant p-value

**Observed:**
- Sample.Type
- location.numeric (look more in detail)
  
**Shannon:**
- Sample.Type
- location.numeric

**Simpson:**
- Sample.Type
- location.numeric

**Chao1:**
- Sample.Type
- location.numeric

**InvSimpson:**
- Sample.Type
- location.numeric


### Linear model for body mass against alpha-diversity

```{r}
results <- glm(Simpson ~ Body.mass..g., data = alpha_div)
anova(results)
summary(results)
```

### Spearman Partial correlation test

```{r}
# Extract all possible covariates
Spearman_covariate <- as.data.frame(alpha_div[, c(3:22, 35)])

# Change Age_category to binary (Juv = 1; Ad = 2)
Spearman_covariate$Age_category <- c(Adult = 2, Juvenile = 1)[Spearman_covariate$Age_category]

# Change Island to binary (Kauai = 1; Bermuda = 2)
Spearman_covariate$Island <- c(Bermuda = 2, Kauai = 1)[Spearman_covariate$Island]

# Change everything to numeric
Spearman_covariate$sex_calc_numeric <- as.numeric(Spearman_covariate$sex_calc_numeric)
Spearman_covariate$location.numeric <- as.numeric(Spearman_covariate$location.numeric)
Spearman_covariate$Island <- as.numeric(Spearman_covariate$Island)
Spearman_covariate$Body.mass..g. <- as.numeric(Spearman_covariate$Body.mass..g.)

Spearman_covariate$Simpson <- as.numeric(Spearman_covariate$Simpson)

Spearman_covariate$PC1 <- as.numeric(Spearman_covariate$PC1)
Spearman_covariate$PC2 <- as.numeric(Spearman_covariate$PC2)
Spearman_covariate$PC3 <- as.numeric(Spearman_covariate$PC3)
Spearman_covariate$PC4 <- as.numeric(Spearman_covariate$PC4)
Spearman_covariate$PC5 <- as.numeric(Spearman_covariate$PC5)
Spearman_covariate$PC6 <- as.numeric(Spearman_covariate$PC6)
Spearman_covariate$PC7 <- as.numeric(Spearman_covariate$PC7)
Spearman_covariate$PC8 <- as.numeric(Spearman_covariate$PC8)
Spearman_covariate$PC9 <- as.numeric(Spearman_covariate$PC9)
Spearman_covariate$PC10 <- as.numeric(Spearman_covariate$PC10)
Spearman_covariate$PC11 <- as.numeric(Spearman_covariate$PC11)
Spearman_covariate$PC12 <- as.numeric(Spearman_covariate$PC12)
Spearman_covariate$PC13 <- as.numeric(Spearman_covariate$PC13)
Spearman_covariate$PC14 <- as.numeric(Spearman_covariate$PC14)
Spearman_covariate$PC15 <- as.numeric(Spearman_covariate$PC15)
Spearman_covariate$PC16 <- as.numeric(Spearman_covariate$PC16)
Spearman_covariate$PC17 <- as.numeric(Spearman_covariate$PC17)
Spearman_covariate$PC18 <- as.numeric(Spearman_covariate$PC18)
Spearman_covariate$PC19 <- as.numeric(Spearman_covariate$PC19)
Spearman_covariate$PC20 <- as.numeric(Spearman_covariate$PC20)
# Remove NA
Spearman_covariate <- Spearman_covariate[!is.na(Spearman_covariate$Body.mass..g.), ]

# Realise a partial Spearman correlation test
Spearman <- pcor(Spearman_covariate, method = "spearman")

# Visualize the correlation matrix
corrplot(Spearman$estimate, method = "number", type = "lower", diag = FALSE)
```

## Beta diversity test

#### Generate distance metrics

```{r, eval = F}
# Set the distance metrics for the matrix
ps_dist_matrix <- phyloseq::distance(ps_Norm, method = "bray")

# Set the distance metrics (Aitchinson) for the matrix
# First convert to CLR
ps_CLR <- microbiome::transform(ps_Norm, "clr")
ps_dist_matrix_clr <- phyloseq::distance(ps_CLR, method = "euclidean")

# Set the distance metrics for the matrix
ps_dist_matrix_unifrac <- phyloseq::distance(ps_Norm, method = "unifrac", weighted = T)
```

### Adonis test

The Adonis test is a version of a PERMANOVA. It is essentially a test for difference in the between beta diversity in the sample variables.

**Significant result**: Indicates difference between samples variables.

#### One Variable Adonis with significant p-value (Bray)

```{r, eval = F}
# Do the Adonis test based on one variables you want
vegan::adonis2(ps_dist_matrix ~ phyloseq::sample_data(ps_Norm)$Age_category, permutations = 999)
```


- Island
- location.numeric
- Sample.Type
- Age_category

#### One Variable Adonis with significant p-value (Aitch)

```{r, eval = F}
# Do the Adonis test for Aitchinson based on one variables you want
vegan::adonis2(ps_dist_matrix_clr ~ phyloseq::sample_data(ps_CLR)$Sample.Type, permutations = 999)
```

- Island
- location.numeric
- Age_category
- Sample.Type

#### One Variable Adonis with significant p-value (Unifrac)

```{r, eval = F}
# Do the Adonis test for unifrac based on one variables you want
vegan::adonis2(ps_dist_matrix_unifrac ~ phyloseq::sample_data(ps_Norm)$Sex, permutations = 999)
```


#### Two or more Variable Adonis with significant p-value (Bray)

```{r, eval = F}
# The Adonis test based on two variables
vegan::adonis2(ps_dist_matrix ~ phyloseq::sample_data(ps_Norm)$Sample.Type + phyloseq::sample_data(ps_Norm)$Island + phyloseq::sample_data(ps_Norm)$Age_category, permutations = 999, by = "terms")
```

- Island/location.numeric but not the opposite
- Age_category still significant with calc_sex
- Sample.Type + location.numeric + Age_category: all significant
- Sample.Type + Island + Age_category: all significant

#### Two or more Variable Adonis with significant p-value (Aitch)

```{r, eval = F}
# The Adonis test based on two variables
vegan::adonis2(ps_dist_matrix_clr ~ phyloseq::sample_data(ps_CLR)$Island + phyloseq::sample_data(ps_CLR)$Sample.Type + phyloseq::sample_data(ps_CLR)$Age_category, permutations = 999, by = "terms")
```

- Island/location.numeric but not the opposite
- Age_category still significant with calc_sex
- Island + Sample_Type + Age_category: all significant
  
```{r, eval = F}
# The Adonis test based on interation of two variables
vegan::adonis2(ps_dist_matrix ~ phyloseq::sample_data(ps_Norm)$Island * phyloseq::sample_data(ps_Norm)$Sex, permutations = 999, by = "terms")
```

# Save the data

```{r, eval = F}
save.image(file.path(save, "Decomtamination_pipeline.RData"))
```